---
title: "Modeling Normal Behavior of generator"
output: pdf_document
html_notebook: default
---

# Prerequesites

The aim of the this notebook is to go through some modeling techniques in order to classify unhealthy behavior of the wind turbines.
The model evaluation technique will be based on the RMSE of the test set. The steps we will go through in this notebook are represented below: 

* Read and split the clean data according to 70-30 ratio (70% for training and 30% for testing)
* Fit a gradient boosting model on the training set and validate the model
* Use evaluation metrics such as RMSE and R-squared to help determine the best model (and do feature selection).
* Once the model is trained and validated, we test the model by predicting the failures on the un-processed data.

- The first approach is to use an isolation forest on a portion of the total data with without any preprocessing. The aim of this approach is to see whether we are able to detect anomalies/failures in the original data set.

- The second approach is to model normal behavior of turbines using the training data from the previous notebook and predict failure values in the test data (that is un-filtered). 
The target variable will be Gen_RPM_Avg and both train and test data will be normalized.

```{r, warning = FALSE, echo = FALSE, include=FALSE}
rm(list=ls())
library(ensaiWind)
library(tidyr)
library(dplyr)
library(readr)
library(lubridate)
library(ggplot2)
library(anomalize)
library(vip)
library(Metrics)
library(xgboost)
library(scales)
library(parallel)
library(gridExtra)
source("functions.R")
```

Recall that the variables we will use for modeling the normal behavior of the generator are :

Gen_RPM_Avg, <- our target
Amb_WindSpeed_Avg,
Prod_LatestAvg_TotActPwr,
Prod_LatestAvg_TotReactPwr,
Nac_Temp_Avg,
Gen_Bear_Temp_Avg,
Gen_Phase1_Temp_Avg,
and Amb_Temp_Avg,
and Blds pitch angle ???? ==> Seems to have a high correlation

RMK : We only choose one Tenperature for the Generator phases, since the correlation between the three are highly correlated.



```{r, include=FALSE}
#-- read the clean data and select gearbox variables
df_gen = read_csv("filtered_data.csv.gz") %>% 
  mutate(Turbine_ID = as.factor(Turbine_ID)) %>%
  select(Datetime,
         Turbine_ID,
         Gen_RPM_Avg,
         Amb_WindSpeed_Avg,
         Prod_LatestAvg_TotActPwr, 
         Prod_LatestAvg_TotReactPwr,
         Nac_Temp_Avg,
         Gen_Bear_Temp_Avg,
         Gen_Phase1_Temp_Avg,
         Amb_Temp_Avg,
         Blds_PitchAngle_Avg) %>%
  filter(Turbine_ID=="T06")

#-- train-validation split
idx_train = seq(1, floor(dim(df_gen)[1]*0.8))
df_gen_train = df_gen[idx_train,]
df_gen_valid = df_gen[-idx_train,]
```

## GAM Model

```{r}
if (! require(gam)) install.packages("gam")
#-- select the training features and target as a matrix and a vector 
X_gen_train = df_gen_train %>%
  select(-Gen_RPM_Avg, -Datetime, -Turbine_ID) %>%
  as.matrix()

Y_gen_train = df_gen_train %>% 
  pull(Gen_RPM_Avg)

#-- select the testing features and target as a matrix and a vector 
X_gen_valid = df_gen_valid %>%
  select(-Gen_RPM_Avg, - Datetime, -Turbine_ID) %>%
  as.matrix()

Y_gen_valid = df_gen_valid %>%
  pull(Gen_RPM_Avg)

#-- train Random forest model

df_gam_train <-data.frame(X_gen_train,Y_gen_train)

gen_gam<-gam(formula = Y_gen_train~s(Amb_WindSpeed_Avg)+
               s(Prod_LatestAvg_TotActPwr)+
               s(Prod_LatestAvg_TotReactPwr)+
               s(Nac_Temp_Avg)+
               s(Gen_Bear_Temp_Avg)+
               s(Gen_Phase1_Temp_Avg)+
               s(Amb_Temp_Avg)+
               s(Blds_PitchAngle_Avg),data = df_gam_train)
```
```{r}
#-- get train prediction and evaluation metrics
Y_gen_train_pred = predict(gen_gam, newdata = df_gam_train)
train_gen_rmse = rmse(Y_gen_train_pred, Y_gen_train)
train_gen_mae = mae(Y_gen_train_pred, Y_gen_train)
train_gen_r2 = cor(Y_gen_train_pred, Y_gen_train)^2

cat("train RMSE is", train_gen_rmse, "\n")
cat("train MAE is", train_gen_mae, "\n")
cat("train R2 is", train_gen_r2, "\n")

df_gam_valid<-data.frame(X_gen_valid,Y_gen_valid)

#-- get test predictions and evaluation metrics
Y_gen_valid_pred = predict(gen_gam, newdata = df_gam_valid)
valid_gen_rmse = rmse(Y_gen_valid_pred, Y_gen_valid)
valid_gen_mae = mae(Y_gen_valid_pred, Y_gen_valid)
valid_gen_r2 = cor(Y_gen_valid_pred, Y_gen_valid)

cat("\nvalidation RMSE is", valid_gen_rmse, "\n")
cat("validation MAE is", valid_gen_mae, "\n")
cat("validation R2 is", valid_gen_r2, "\n")
```


## Random Forest model

```{r}
library(ranger)
#-- select the training features and target as a matrix and a vector 
X_gen_train = df_gen_train %>%
  select(-Gen_RPM_Avg, -Datetime, -Turbine_ID) %>%
  as.matrix()

Y_gen_train = df_gen_train %>% 
  pull(Gen_RPM_Avg)

#-- select the testing features and target as a matrix and a vector 
X_gen_valid = df_gen_valid %>%
  select(-Gen_RPM_Avg, - Datetime, -Turbine_ID) %>%
  as.matrix()

Y_gen_valid = df_gen_valid %>%
  pull(Gen_RPM_Avg)

#-- train Random forest model

df_ranger_train <-data.frame(X_gen_train,Y_gen_train)

gen_ranger<-ranger(formula = Y_gen_train~.,
            data = df_ranger_train,
            num.trees = 500,
            verbose = FALSE)

```

```{r}
#-- get train prediction and evaluation metrics
Y_gen_train_pred = predict(gen_ranger, data = df_ranger_train)$predictions
train_gen_rmse = rmse(Y_gen_train_pred, Y_gen_train)
train_gen_mae = mae(Y_gen_train_pred, Y_gen_train)
train_gen_r2 = cor(Y_gen_train_pred, Y_gen_train)^2

cat("train RMSE is", train_gen_rmse, "\n")
cat("train MAE is", train_gen_mae, "\n")
cat("train R2 is", train_gen_r2, "\n")

df_ranger_valid<-data.frame(X_gen_valid,Y_gen_valid)

#-- get test predictions and evaluation metrics
Y_gen_valid_pred = predict(gen_ranger, data = df_ranger_valid)$predictions
valid_gen_rmse = rmse(Y_gen_valid_pred, Y_gen_valid)
valid_gen_mae = mae(Y_gen_valid_pred, Y_gen_valid)
valid_gen_r2 = cor(Y_gen_valid_pred, Y_gen_valid)

cat("\nvalidation RMSE is", valid_gen_rmse, "\n")
cat("validation MAE is", valid_gen_mae, "\n")
cat("validation R2 is", valid_gen_r2, "\n")
```


## XGBOOST Model

We train our xgboost model :

```{r}
#-- select the training features and target as a matrix and a vector 
X_gen_train = df_gen_train %>%
  select(-Gen_RPM_Avg, -Datetime, -Turbine_ID) %>%
  as.matrix()

Y_gen_train = df_gen_train %>% 
  pull(Gen_RPM_Avg)

#-- select the testing features and target as a matrix and a vector 
X_gen_valid = df_gen_valid %>%
  select(-Gen_RPM_Avg, - Datetime, -Turbine_ID) %>%
  as.matrix()

Y_gen_valid = df_gen_valid %>%
  pull(Gen_RPM_Avg)

#-- train xgboost model
gen_xgb = xgboost(
  data = X_gen_train,
  label = Y_gen_train,
  verbose = 0,
  nrounds = 2000,                 #-- max number of boosting iterations
  params = list (eta = 0.2,       #-- learning rate for the gradient descent step
                 nthread = 10)   #-- for parallel computing on 10 cores
)
```


```{r}
#-- get train prediction and evaluation metrics
Y_gen_train_pred = predict(gen_xgb, newdata = X_gen_train)
train_gen_rmse = rmse(Y_gen_train_pred, Y_gen_train)
train_gen_mae = mae(Y_gen_train_pred, Y_gen_train)
train_gen_r2 = cor(Y_gen_train_pred, Y_gen_train)^2

cat("train RMSE is", train_gen_rmse, "\n")
cat("train MAE is", train_gen_mae, "\n")
cat("train R2 is", train_gen_r2, "\n")

#-- get test predictions and evaluation metrics
Y_gen_valid_pred = predict(gen_xgb, newdata = X_gen_valid)
valid_gen_rmse = rmse(Y_gen_valid_pred, Y_gen_valid)
valid_gen_mae = mae(Y_gen_valid_pred, Y_gen_valid)
valid_gen_r2 = cor(Y_gen_valid_pred, Y_gen_valid)

cat("\nvalidation RMSE is", valid_gen_rmse, "\n")
cat("validation MAE is", valid_gen_mae, "\n")
cat("validation R2 is", valid_gen_r2, "\n")
```


```{r, fig.width=12, fig.height=6, warning=FALSE}
#-- build data frames of predicted and observed values of train and test data
g1 = data.frame(observed = df_gen_valid$Gen_RPM_Avg, predicted = Y_gen_valid_pred)
g2 = data.frame(observed = df_gen_train$Gen_RPM_Avg, predicted = Y_gen_train_pred)

#write.csv(g1, file=gzfile("data_valid_gen.csv.gz"), row.names = FALSE)
#write.csv(g2, file=gzfile("data_train_gen.csv.gz"), row.names = FALSE)

#-- plot predicted values against real values and see if its linear
p1 = ggplot(g1, aes(x=predicted, y=observed)) + 
  geom_point(alpha = 0.05) + theme_bw() +
  ggtitle("Observed vs predicted values in validation set Gen_RPM_Avg") + 
  geom_smooth(method='lm', formula = y~x, color = "steelblue") +
  geom_line(aes(y = predicted), size = 1, color = "red", linetype = "dashed")

p2 = ggplot(g2, aes(x=predicted, y=observed)) + 
  geom_point(alpha = 0.05) + theme_bw() +
  ggtitle("Observed vs predicted values in training set Gen_RPM_Avg")+ 
  geom_smooth(method='lm', formula = y~x, color = "steelblue") +                        #-- the linear regression line
  geom_abline(intercept = 0, slope = 1, size = 1, color = "red", linetype = "dashed") + #-- the slope line (y=x)
  scale_color_discrete(name = "Type", labels = c("lm line", "slope"))

grid.arrange(p1, p2, ncol=2)
```


CHECK FOR LOW VALUES (IF NO WIND, COULD BE REMOVED)

To get a feature importance plot, we do the following :

```{r, fig.width=10, fig.height=4, warning=FALSE, fig.align='center'}
vip::vip(gen_xgb) + theme_bw() + ggtitle("Feature importance of the generator")
```

We can also plot the residual's histogram for the training set :

```{r}
res_train=Y_gen_train_pred - Y_gen_train
res_valid=Y_gen_valid_pred - Y_gen_valid
hist(res_train,xlim = c(-30,30),freq = FALSE)
quantile(res_valid,probs=c(0.025,0.975))
sd(res_valid)
sd(res_train)
```


## Normal behavior boundaries according to the threshold

To test the algorithm, we will predict an abnormal behavior on the test data (the last 6 months of data) according to the fixed threshold above

```{r, include=FALSE}
#-- read the unfiltered data and select gearbox variables
df_gen_test = read_csv("unfiltered_data.csv.gz") %>% 
  mutate(Turbine_ID = as.factor(Turbine_ID)) %>%
  select(Datetime,
         Turbine_ID,
         Gen_RPM_Avg,
         Amb_WindSpeed_Avg,
         Prod_LatestAvg_TotActPwr, 
         Prod_LatestAvg_TotReactPwr,
         Nac_Temp_Avg,
         Gen_Bear_Temp_Avg,
         Gen_Phase1_Temp_Avg,
         Amb_Temp_Avg,
         Blds_PitchAngle_Avg) %>%
  filter(Turbine_ID=="T06") %>%
  as_tibble() 

#-- get the test features and the test target as a matrix and a vector
X_gen_test = df_gen_test %>% 
  select(-Datetime, -Gen_RPM_Avg, -Turbine_ID) %>%
  as.matrix()

Y_gen_test = df_gen_test %>%
  pull(Gen_RPM_Avg)
```

```{r}
#-- get predictions and compute evaluation metrics
Y_gen_test_pred = predict(gen_xgb, newdata = X_gen_test)
test_gen_rmse = rmse(Y_gen_test_pred, Y_gen_test)
test_gen_mae = mae(Y_gen_test_pred, Y_gen_test)
test_gen_r2 = cor(Y_gen_test_pred, Y_gen_test)

cat("Test RMSE is", test_gen_rmse, "\n")
cat("Test MAE is", test_gen_mae, "\n")
cat("Test R2 is", test_gen_r2, "\n")
```


We visualize the results below and add lines according to the thresholds we fixed to represent the normal behavior boundaries

```{r, fig.width=8, fig.height=6, warning=FALSE}
#-- build data frames of predicted and observed values of train and test data
g3 = data.frame(observed = df_gen_test$Gen_RPM_Avg, predicted = Y_gen_test_pred)
#write.csv(g3, file=gzfile("data_test_gen.csv.gz"), row.names = FALSE)

#-- plot predicted values against real values and see if its linear
ggplot(g3, aes(x=predicted, y=observed)) + 
  geom_point(alpha = 0.05) + theme_bw() +
  ggtitle("Observed vs predicted values in validation set Gen_RPM_avg")
```


```{r}
thresh=50
df_gen_pred = df_gen_test %>%
  mutate(Predict_RPM = Y_gen_test_pred,
         Max_Diff = abs(Predict_RPM - Gen_RPM_Avg))

df_gen_pred_thresh = df_gen_pred %>%
  filter(Max_Diff>thresh)

anomalies_gen<-df_gen_pred$Datetime[which(df_gen_pred$Max_Diff >thresh)]

cat(dim(df_gen_pred_thresh)[1], "anomalies have been predicted using the threshold method")

#write.csv(df_gen_pred_thresh, file=gzfile("data_anomalies_gen.csv.gz"), row.names = FALSE)
```

We can plot the anomalies predicted as follow:

```{r, fig.width=12, fig.height=5, fig.align='center'}
p2 = df_gen_pred_thresh %>%
  #select(Amb_WindSpeed_Avg,Gen_RPM_Avg,Predict_RPM, Datetime, -Amb_Temp_Avg) %>%
  select(Amb_WindSpeed_Avg,Datetime) %>%
  gather(key = "Variables", value = "values", -Datetime) %>%
  ggplot(aes(x=Datetime, y=values, color = Variables)) + 
  geom_point(alpha = 0.5) + ggtitle("Anomalies predicted by our threshold method") + theme_bw() + 
  scale_x_datetime(date_labels = "%Y-%m-%d", breaks = date_breaks("1 month"))
  #geom_vline(xintercept=as.POSIXct("11-08-2017",format="%d-%m-%Y"), linetype='dashed', color='blue', size=1)+
  #geom_vline(xintercept=as.POSIXct("19-08-2017",format="%d-%m-%Y"), linetype='dashed', color='blue', size=1)+
  #geom_vline(xintercept=as.POSIXct("20-08-2017",format="%d-%m-%Y"), linetype='dashed', color='blue', size=1)+
  #geom_vline(xintercept=as.POSIXct("21-08-2017",format="%d-%m-%Y"), linetype='dashed', color='blue', size=1)
  #geom_vline(xintercept=as.POSIXct("12-09-2017",format="%d-%m-%Y"), linetype='dashed', color='blue', size=1)+
  #geom_vline(xintercept=as.POSIXct("16-09-2017",format="%d-%m-%Y"), linetype='dashed', color='blue', size=1)+
  #geom_vline(xintercept=as.POSIXct("18-10-2017",format="%d-%m-%Y"), linetype='dashed', color='blue', size=1)+
  #geom_vline(xintercept=as.POSIXct("19-10-2017",format="%d-%m-%Y"), linetype='dashed', color='blue', size=1)
  
p2


```

## Detect anomalies using the TS decomposition

If we apply the time series decomposition method on the original data (to maintain the structure of all 2 years of data if there is any seasonality envolved) and get the anomalies that only happened in dates of the test data, we find the following:

```{r}
#-- load original data
data_total = read_csv("original_data.csv.gz") %>% 
  mutate(Turbine_ID = as.factor(Turbine_ID)) %>%
  filter(Turbine_ID == "T01") %>%
  droplevels() %>%
  as_tibble()

#-- display anomalies using the TS decomposition on all 2 years of data and filter only the test dates
d = data_total %>%
  find_my_anomalies_ts() #%>%
  #filter(Datetime > as.POSIXct("2017-07-02 11:10:00",format="%Y-%m-%d %H:%M:%OS"))    #-- to get only the dates of the test set

cat(dim(d)[1], "anomalies have been found in the test set")
```

We found 5427 anomalies using the Times Series decomposition. We can visualize these points as follows :

```{r, fig.width=12, fig.height=8, fig.align='center'}
p1 = d %>%
  select(contains("Temp"), Datetime, -Amb_Temp_Avg) %>%
  gather(key = "Variables", value = "values", -Datetime) %>% 
  ggplot(aes(x=Datetime, y=values, color = Variables)) + 
  geom_point(alpha = 0.5) + ggtitle("Anomalies detected by the time series decomposition method") + theme_bw() + 
  scale_x_datetime(date_labels = "%Y-%m-%d", breaks = date_breaks("1 month"))

grid.arrange(p1,p2)
```

TO DO : Cross check anomalies detected by TS decomposition with the ones detected by our threshold method.


We can plot the actual values predicted in time along with the observed values in the test set.

```{r}
df_gen_pred_thresh %>%
  select(Gen_RPM_Avg,Predict_RPM, Datetime, -Amb_Temp_Avg) %>%
  gather(key = "Variables", value = "values", -Datetime) %>%
  ggplot(aes(x=Datetime, y=values, color = Variables)) +
  geom_point(alpha = 0.5, size = 0.7) + 
  #ggtitle("Anomalies predicted by our threshold method") + 
  theme_bw() +
  scale_x_datetime(date_labels = "%Y-%m-%d", breaks = date_breaks("1 month")) +
  theme(plot.title = element_text(size = 10, face = "bold", hjust = 0.6), 
        legend.title = element_text(size = 8, face = "bold"),
        legend.text = element_text(size = 6),
        axis.text=element_text(size=6),
        axis.title=element_text(size=8,face="bold"))
```
